---
title: map(expr) + renderUI -> dynamic shiny
author: Tom Greenwood
date: '2019-04-15'
slug: map-expr-renderui-dynamic-shiny
categories:
  - shiny
  - R
  - tidy eval
tags:
  - nse
  - shiny
---

I wrote my first shiny widget. A shiny widget is an html widget which uses shiny functions. In form it resembles a shiny app very closely, only the shinyApp function is wrapped in another function, which can take whatever arguments you like. More background can be found [here](https://bookdown.org/yihui/rmarkdown/shiny-widgets.html). 

In order for the shiny widget to run, it needs a shiny runtime behind the document, which means it needs a server, just like a shiny app. For that reason, the widget it's running in this actual document, it's running on at [shinyapps.io](https://greenwood.shinyapps.io/filter_widget/#). I've embedded it below. 


<iframe width="700" height="400" scrolling="yes" frameborder="no" src="https://greenwood.shinyapps.io/filter_widget/"></iframe>

Just a bit of explanation about how it works: The key is in mapping `checkboxGroupInput` (or some other input function) across the names of the `data.frame` passed to the widget. A slightly simpler version would be

```{r warning=FALSE, message=FALSE}
library(shiny)
library(tidyverse, warn.conflicts = FALSE)

# invent some names
c("john", "paul", "ringo") %>%
# map a suitable shiny input across them
map(. %>% { expr(checkboxInput(!!., !!. )) } %>%  eval ) %>% 
tagList 

  
  
```

Let's break this down one step further. First, why use tagList? Well, because `map` returns a list, and we don't want a list, we wants legit html. To see the hmtl in markdown without it rendering need a bit of a workaround, so the next code chunk isn't actually run, but if you run

```{r eval=FALSE}
checkboxInput("john", "john")
```

In an R console, what you get is

```html
<div class="form-group shiny-input-container">
  <div class="checkbox">
    <label>
      <input id="john" type="checkbox"/>
      <span>john</span>
    </label>
  </div>
</div>
```

I know I don't actually use tagList in the widget, I use renderUI, but they're performing the same role there. In the shinywidget it needs renderUI, but in markdown, outside of a widget you could tagList.

So, mapping `checkboxInput` across a list of names gives you a list of chunks of html. `tagList` just combines a list of chunks of html into one big chunk, without it, you just get the html from the last element of the list (I'm not sure why). 

```{r}
list(checkboxInput("john", "john"), 
        checkboxInput("paul", "paul"),
        checkboxInput("ringo", "ringo"))
```

v.s.

```{r}
tagList(checkboxInput("john", "john"), 
        checkboxInput("paul", "paul"),
        checkboxInput("ringo", "ringo"))
```

And finally, good ol' map(expr), which makes it easy to create any list of expressions you want.

```{r}
letters[1:5] %>% map(~ expr(selectInput(!!., !!., choices = c(TRUE, FALSE))))
```

And that's it!