---
title: visualising models in 3d with plotly and nse functions
author: Tom Greenwood
date: '2019-02-09'
slug: visualising-models-in-3d-with-plotly-and-nse-functions
categories:
  - R
tags:
  - functions
  - plotly
  - visualisation
---



One of the areas I'm planning to improve on is visualisation. Ultimately, this means D3, but in the meantime I wrote a fairly useful function for showing the fit of a model with two inputs, along with the data, using plotly. First, load the necessary libraries:
```{r echo=TRUE, message=FALSE}
library(plotly)
library(randomForest)
library(rlang)
library(dplyr)
library(magrittr)
```
(you could actually write thing function without rlang or magrittr, but it's just a bit neater with them.) Here's the function:
```{r}
plot_3d <- function(df, x, y, z) {
  
  x <- enquo(x)
  y <- enquo(y)
  z <- enquo(z) 
  
  # build a model 
  m <- randomForest(expr(!!get_expr(x) ~ !!get_expr(y) + !!get_expr(z)) %>% as.formula, df)
  
  # add the predictions to df
  df %<>% mutate(preds = predict(m, newdata = df ))
  
  plot_ly(df) %>%
    add_trace(x = x, y = y, z = z, 
                opacity = 0.7,
                type = "scatter3d", 
                mode = "markers",
                colors = c("grey", "blue"),
                name = "true values",
                size = 10
              ) %>%
    add_trace(x = ~preds, y = y, z = z,
                opacity = 0.7,
                type = "scatter3d", 
                mode = "markers",
                name = "modelled values",
                size = 10
                ) %>%
    layout(scene = list(xaxis = list(title = quo_text(x)),
                        yaxis = list(title = quo_text(y)),
                        zaxis = list(title = quo_text(z))
    )
    )
  
}

economics %>% plot_3d(pce, psavert, uempmed)
```
Look's like random forrest models are pretty decent for non-linear regression!


