---
title: running rust in .Rmd
author: Tom
date: '2019-12-22'
slug: running-rust-in-rmd
categories:
  - rust
  - knitr
tags: []
---

I have been wanting to learn Rust for some time, but one of the things that's been on my mind is: If I learn some good Rust, how am I going to blog about it? This site is build on blogdown, which is R markdown + Hugo, so any code is blog posts needs to be run by knitr (which is the package with runs lots of different languages and outputs the result in rmarkdown). 

I have thought about switching to some other form of multi-language notebook, but as far as I'm aware there's nothing with as wide a range of supported languages as knitr. Please let me know in the comments if there is!

Anyway, knitr supports building your own language engines, so I did one for Rust.

```{r }

wd_tempfile = function(...) basename(tempfile(tmpdir = '.', ...))

eng_rust = function(options) {
  f = wd_tempfile('main', '.rs')
  xfun::write_utf8(code <- options$code, f)
  on.exit(unlink(f), add = TRUE)
 
  cmd = '/home/tom/.cargo/bin/rustc'
  
  rustc_args = sprintf(' %s', f)
  
  message('running: ', cmd, rustc_args)
  tryCatch(
    system2(cmd, rustc_args, stdout = TRUE, stderr = TRUE, env = options$engine.env),
    error = function(e) {
      if (!options$error) stop(e)
    }
  )

  exef = sub('.rs', '', f)
  run_cmd = paste('./', exef, sep = '')

  extra = if (options$eval) {
    message('running: ', run_cmd)
    tryCatch(
      system2(run_cmd, stdout = TRUE, stderr = TRUE, env = options$engine.env),
      error = function(e) {
        if (!options$error) stop(e)
        'Error in executing rust code'
      }
    )
  }

  if (options$results == 'hide') extra = NULL

  knitr::engine_output(options, code, extra)  
}

knitr::knit_engines$set(rust = eng_rust)

```

It works

```{rust}

fn main() {
    println!("Hello, world!");
}
```

It also works with the standard library (as you'd expect). 

```{rust}
use std::iter;

fn main() {
    let v = vec![1, 2, 3, 4, 5];
    v.iter().for_each(|x| println!("{}", x));
}
```

However, it wouldn't work with additional Rust dependiencies. Cargo is the dependency management tool for Rust. So, I tried a different engine configuration, using Cargo. 

```{r}

eng_rust = function(options) {
  path = 'src/main.rs'
  xfun::write_utf8(code <- options$code, path)
  on.exit(unlink(path), add = TRUE)
 
  cmd = '/home/tom/.cargo/bin/cargo'
  
  extra = if (options$eval) {
    message('running: ', cmd)
    tryCatch(
      system2(cmd, 'run', stdout = TRUE, stderr = TRUE, env = options$engine.env),
      error = function(e) {
        if (!options$error) stop(e)
        'Error in executing rust code'
      }
    )
  }

  if (options$results == 'hide') extra = NULL

  knitr::engine_output(options, code, extra)  
}

knitr::knit_engines$set(rust = eng_rust)
```

With this, I can import an external library from 'crates' and it'll still run nicely in r markdown. This seems useful.

```{rust}
use ferris_says::say; // from the previous step
use std::io::{stdout, BufWriter};

fn main() {
    let stdout = stdout();
    let out = b"Hello fellow Rustaceans!";
    let width = 24;

    let mut writer = BufWriter::new(stdout.lock());
    say(out, width, &mut writer).unwrap();
}

```

Some caveats:

1. I have hard coded the paths to rustc and cargo as they are on my machine, so this isn't fit to be a generic engine for the knitr package just yet.
2. In the first engine I assumed that the machine running the code is using a linux OS. `./file` isn't going to work on a Windows machine.
3. In the second engine `src/` and `Cargo.toml` (with the necessary dependencies listed) have to be in the same directory as the markdown being run. Again, this works for me, but might cause issues as part of a package, because that directory structure might not always be possbile.

TODOS: 

1. Make at least one of the engines generic enough to submit a pull request to knitr.
2. Do something interesting with Rust.
